Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-12 Manuscript/manuscript.tex
Update text

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_na_k_cl_wat_phos_models-compar_5-6_NaCl-and-KCl-series.pdf}
  \caption{\label{fig:nacl-dens_PCPS} 
    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} and \ce{Cl^-} along the membrane normal axis 
    for the negatively charged membranes with the composition of 5\,PC:1\,PS using ECC-lipids and ECC-ions
    from simulations with added concentrations of \ce{KCl} or \ce{NaCl}.  
    The top    profile shows the simulation with an additional \ce{KCl} concentration and \ce{Na^+} counterions. 
    The bottom profile shows the simulation with an additional \ce{NaCl} concentration and \ce{Na^+} counterions, which are not distinguished from the added salt. 
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
  }
  \todo{I am not sure if we need this figure at all.}
\end{figure}

%SAMULI: I WOULD LEAVE THIS PART OUT
%% This§: 
%% adsorption and affinities of Na and K to PC:PS memb
%% from densities and rel.surf.excess
%% 
%%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
%%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
%%
%The affinities of \ce{Na^+} and \ce{K^+} to phospholipid membranes
%can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$. 
%Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
%without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
%Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
%This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
%showing that the density profile of \ce{K^+} along the membrane normal 
%decays clearly faster towards the membrane centre compared to the density of water. 
%In contrast, the density profile of \ce{Na^+} has a shoulder 
%at the region of the phosphate groups,
%which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
%by yielding a small positive value of $\Gamma^{w}_{Na}$.

@@

The headgroup order parameters of PC lipids decrease more upon addition of \ce{CaCl2}
in POPC:POPS (5:1) mixtures than in pure PC bilayers due to the higher calcium binding
affinity to the negatively charged lipid bilayers \cite{akutsu81,altenbach84,roux90,nmrlipids_proj4}.
Following our previous work \cite{nmrlipids_proj4}, we use the experimental
headgroup order parameter data as a function of CaCl$_2$ concentration
from POPC:POPS (5:1) mixture to evaluate the calcium binding affinity to lipid bilayers containing PS lipids.
The decrease of the PC headgroup order parameters upon addition of \ce{CaCl2} 
is overestimated in the Lipid17/Dang simulation, while ECC-POPS simulations give results
closer to the experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
This indicates that the lower binding affinity of calcium in ECC-POPS
simulation is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
However, the steep decrease of the order parameters at low calcium concentrations
observed especially in the negatively charged bilayers is not captured by the ECC-lipid simulations.

@@

The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
are very challenging to interpret from both experiments \cite{roux90} 
and also from MD simulations \cite{nmrlipids_proj4}. 
The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
are not found to follow the experimental profiles, 
nor any common trend,
with the currently available models \cite{nmrlipids_proj4}. 
In Fig.~\ref{fig:delta_ordPar_CaCl_PCPS},
we plot these structural parameters for one of the models in that work, Lipid17, 
and for ECC-lipids. 
Although the agreement between ECC-lipids and experiments is not optimal,
the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
are well captured by the model. 
For example the order parameter $\beta$ in POPS 
adopts a saturated value in experiments and ECC-lipids,
but it is growing with the \ce{CaCl2} concentration in Lipid17.
In addition, 
the forking of the order parameters $\alpha$ 
is almost constant in the experiment and ECC-lipids,
but it varies greatly in th Lipid17.

The source of this discrepancy may arise from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
from the sub-optimal structure of the lipids (Fig.~\ref{simVSexpNOions_POPS},
and also possibly from the approximate mean-field treatment of the electronic polarization.

It has been pointed out in previous experimental works
that the conformational flexibility of PS head groups 
is different from other lipids \cite{browning80}. 
While the static shielding tensor of PS is not substantially different from PC or PE, 
the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
It is suggested that the PS head group is more rigid compared to the other phospholipids, 
which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.

@@

\begin{table}[tb!] 
\centering
  \caption{Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
           relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
           and percentages of the population 
           of bound Ca$^{2+}$ to various moieties 
           in a neutral membrane composed of POPC
           and in a negatively charged membrane with the composition 5\,PC:1\,PS.
           \label{tab:binding}} 
  \begin{tabular}{ l | c c } 
	                     &  5\,POPC:1\,POPS &  POPC   \\
	\hline
	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
	\hline
                             &  \multicolumn{2}{c}{ } \\
        interacting moiety   &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
	\hline
	     PC              &   59   &  100   \\
	     PO$_4$    in PC &   41   &   67   \\
	     carbonyls in PC &   <1   &   ~1   \\
	\hline
	     PS              &    8   &        \\ 
	     PO$_4$  in PS   &    2   &        \\
	     COO$^-$ in PS   &    4   &        \\
	     carbonyls in PS &   <1   &        \\
	\hline
	both PC and PS       &   33   &        \\
  \end{tabular} 
\end{table}

% Gamma and higher affinity of PS to CA2+
The higher affinity of calcium to PS head groups compared to PC
is well demonstrated using the relative surface excess, $\Gamma ^w _{Ca}$,
summarized in Table~\ref{tab:binding}. 
While $\Gamma^{w}_{Ca}$ at a neutral POPC bilayer
is       $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$      \cite{melcr18}, 
its value for the PC:PS (5:1) bilayer is $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$
at comparable buffer concentrations  $350 \, \mathrm{mM}$ resp. $400\, \mathrm{mM}$.

>>>>

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/density_profiles_na_k_cl_wat_phos_models-compar_5-6_NaCl-and-KCl-series.pdf}
  \caption{\label{fig:nacl-dens_PCPS} 
    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} and \ce{Cl^-} along the membrane normal axis 
    for the negatively charged membranes with the composition of 5\,PC:1\,PS using ECC-lipids and ECC-ions
    from simulations with added concentrations of \ce{KCl} or \ce{NaCl}.  
    The top    profile shows the simulation with an additional \ce{KCl} concentration and \ce{Na^+} counterions. 
    The bottom profile shows the simulation with an additional \ce{NaCl} concentration and \ce{Na^+} counterions, which are not distinguished from the added salt. 
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
  }
  \todo{I am not sure if we need this figure at all or should we move it to SI.}
  \todo{The concentrations in the figure do not match with the table.}
\end{figure}

% This§: 
% adsorption and affinities of Na and K to PC:PS memb
% from densities and rel.surf.excess
% 
%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
%
The influence of negatively charged PS on the counterion binding affinity can
be estimated by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
between POPC:POPS (5:1) mixture and pure POPC \cite{melcr18}.
%Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
%without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure POPC, calculated from
ECC-POPC simulation, was reported in the previous work \cite{melcr18}. Here, the ECC-lipid simulations
give
%$\Gamma ^{w} _{\rm K}=-0.12 \pm ??$ nm$^{-2}$ and
$\Gamma ^{w} _{\rm Na}=0.09 \pm ??$ nm$^{-2}$ for 0.621 M sodium in pure POPC:POPS (5:1) mixture.
Therefore, the presence of $\sim$17\% POPS in POPC bilayer increases the relative surface
excess of sodium with respect to water from negative to positive. As discussed above,
the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
ECC-lipid simulations. However, the difference may not be affected because similar small
overestimation seem to occur also in pure POPC simulation where the headgroup order parameter
response to NaCl is also slightly overestimated \cite{meclr18}.
%Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
%This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
%showing that the density profile of \ce{K^+} along the membrane normal 
%decays clearly faster towards the membrane centre compared to the density of water. 
%In contrast, the density profile of \ce{Na^+} has a shoulder 
%at the region of the phosphate groups,
%which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
%by yielding a small positive value of $\Gamma^{w}_{Na}$.

@@

The headgroup order parameters of PC lipids decrease more upon addition of \ce{CaCl2}
in POPC:POPS (5:1) mixtures than in pure PC bilayers due to the higher calcium binding
affinity to the negatively charged lipid bilayers \cite{akutsu81,altenbach84,roux90,nmrlipids_proj4}.
Following our previous work \cite{nmrlipids_proj4}, we use the experimental
headgroup order parameter data as a function of CaCl$_2$ concentration
from POPC:POPS (5:1) mixture to evaluate the calcium binding affinity to lipid bilayers containing PS lipids.
The decrease of the PC headgroup order parameters upon addition of \ce{CaCl2} 
is overestimated in the Lipid17/Dang simulation, while ECC-POPS simulations give results
closer to the experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
This indicates that the lower binding affinity of calcium in ECC-POPS
simulation is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
The relative surface excess of calcium with respect to water, $\Gamma ^w _{Ca}$,
calculated from ECC-lipid simulation gives $\Gamma^{w}_{Ca} = 0.24\mathrm{nm^{-2}}$ for the PC:PS (5:1) bilayer with 409 mM CaCl$_2$
and $\Gamma^{w}_{Ca} = 0.06\mathrm{nm^{-2}}$ for pure POPC bilayer with 350 mM CaCl$_2$ \cite{melcr18},
illustrating the higher affinity of calcium to bilayers containing PS lipids.
However, the steep decrease of the order parameters at low calcium concentrations
observed especially in the negatively charged bilayers is not captured by the ECC-lipid simulations.

@@

Experimentally measured headgroup order parameters of POPS exhibit a strong dependence
on low CaCl$_2$ concentrations with a rapid saturation below 100 mM \cite{roux90}.
The changes are again overestimated by Lipid17/Dang simulations, which
also exhibits non-systematic changes in both PS headroup orientation and order parameters
upon addition of CaCl$_2$ (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
In ECC-lipid simulations,  the P--N and C$_{\beta}$--C$_{\gamma}$ vectors systematically
tilt more perpendicular and parallel to the membrane interface, respectively,
and the changes of the PS headgroup order parameters are again closer to the
experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).
Despite significant improvements in the ECC-lipid against Lipid17/Dang
and other force fields for the response of PS headgroup order parameters to
the bound calcium \cite{nmrlipids_proj4}, the strong dependence of both PC and PS
headgroup order parameters on low concetrations of CaCl$_2$ is not reproduced by
the model (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). In addition to the inaccurate
interactions between calcium ions and PS headgroup, the potential sources of this
discrepancy include the above observed slight overestimation of \ce{Na^+} counterions and
inaccurate structures of the lipid headgroup (Fig.~\ref{simVSexpNOions_POPS}).
%and also possibly from the approximate mean-field treatment of the electronic polarization.

%are again closer to the experiments is not optimal,
%the general trends and behaviour even at highly concentrated solutions of \ce{CaCl2} 
%are well captured by the model. 
%For example the order parameter $\beta$ in POPS 
%adopts a saturated value in experiments and ECC-lipids,
%but it is growing with the \ce{CaCl2} concentration in Lipid17.
%In addition, 
%the forking of the order parameters $\alpha$ 
%is almost constant in the experiment and ECC-lipids,
%but it varies greatly in th Lipid17.

%The structural changes in POPS induced by increasing concentrations of \ce{CaCl2} 
%are very challenging to interpret from both experiments \cite{roux90} 
%and also from MD simulations \cite{nmrlipids_proj4}. 
%The changes of the order parameters $S^\alpha$ and $S^\beta$ in PS head groups from simulations 
%are not found to follow the experimental profiles, 
%nor any common trend, with the currently available models \cite{nmrlipids_proj4}.

%SAMULI: We will not go to the spin relaxation time analysis in this work
%and average angles do not give information about dynamics
%
%It has been pointed out in previous experimental works
%that the conformational flexibility of PS head groups 
%is different from other lipids \cite{browning80}. 
%While the static shielding tensor of PS is not substantially different from PC or PE, 
%the $^{31}$P~$T_1$ relaxation times are significantly shorter for PS. 
%It is suggested that the PS head group is more rigid compared to the other phospholipids, 
%which is also corroborated by a larger chemical shielding anisotropy of PS \cite{browning80}.

@@

\begin{table}[tb!] 
\centering
  \caption{%Bulk concentrations, $C _{Ca}$, and buffer concentrations, $C' _{Ca}$, of Ca$^{2+}$;
           %relative surface excess of calcium with respect to water ($\Gamma_{Ca}^{\rm water}$); 
    %and
    Percentages of the population 
           of bound Ca$^{2+}$ to various moieties 
           in a neutral membrane composed of POPC
           and in a negatively charged membrane with the composition 5\,PC:1\,PS.
           \label{tab:binding}} 
  \begin{tabular}{ l | c c } 
        interacting moiety    &  \multicolumn{2}{c}{percentage of bound \ce{Ca^{2+}} } \\
%	\hline
%	$C _{Ca}\,/\,\mathrm{mM}$  &  $240\pm 10 $  &  $280\pm 10 $  \\
%	$C'_{Ca}\,/\,\mathrm{mM}$  &  $400\pm 10 $  &  $350\pm 10 $  \\
%	$\Gamma_{Ca}^{\rm water}\, / \,\mathrm{nm}^{-2}$  &  $0.24 \pm 0.01 $  &  $0.06 \pm 0.01 $  \\
%	\hline
%                             &  \multicolumn{2}{c}{ } \\
                             &  5\,POPC:1\,POPS &  POPC   \\
	\hline
	     PC              &   59   &  100   \\
	     PO$_4$    in PC &   41   &   67   \\
	     carbonyls in PC &   <1   &   ~1   \\
	\hline
	     PS              &    8   &        \\ 
	     PO$_4$  in PS   &    2   &        \\
	     COO$^-$ in PS   &    4   &        \\
	     carbonyls in PS &   <1   &        \\
	\hline
	both PC and PS       &   33   &        \\
  \end{tabular} \\
  \todo{I have commented out the surface excess values (available in text) and concentrations.}
\end{table}
